---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document: 
    keep_md: yes
---

Set future code chunks to always be visible :

```{r setoptions, echo=TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data

### Raw data

Read and store data :
```{r getdata}
activitydf <- read.csv(unzip("activity.zip"))
```


Preview of raw data :
```{r preview.rawdata}
str(activitydf)
head(activitydf, n =3)
tail(activitydf, n =3)
```


### Formated data

Format into date type to the concerned variable :
```{r date.format, message=FALSE}
library(dplyr)
activitydf$date <- strptime(activitydf$date, "%Y-%m-%d")
str(activitydf)
```


## What is mean total number of steps taken per day?

Histogram of the total number of steps taken each day :
```{r totalstep.hist, message = FALSE}
temp <- activitydf %>% group_by(date) %>%
      summarise(stepsperday = sum(steps, na.rm = TRUE))
hist(temp$stepsperday,
     xlab = "Steps per day",
     main = "Numbers of days for a certain amount of steps per day")
```


Mean and median of the total number of steps taken per day :
```{r mean.median, message = FALSE}
activitydf %>% group_by(date) %>%
      summarise(mean.stepsperday = mean(steps),
                median.stepsperday = median(steps)) %>%
      as.data.frame %>% head
```


## What is the average daily activity pattern?

Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days :
```{r interval.timeplot, message = FALSE}
activitydf %>% group_by(interval) %>% 
      summarise(mean(steps, na.rm = TRUE)) %>%
      plot(type = 'l', col = "#53979E", lwd = 2,
           xlab = "5-minutes interval [HoursMinutes]", ylab = "Average number of steps",
           main = "Average number of steps taken each day by 5-minutes day interval")
```


5-minute interval, on average across all the days in the dataset, containing the maximum number of steps :
```{r maxinterval, message = FALSE}
temp <- activitydf %>% group_by(interval) %>% 
      summarise(avg = mean(steps, na.rm = TRUE))

temp$interval[which(max(temp$avg) == temp$avg)]
```


## Imputing missing values

Number of NAs in the dataset :
```{r na}
sum(is.na(activitydf))
```


Filling missing "steps" values with median "steps" calculated for all days by "interval" :
```{r na.fill}
filt <- is.na(activitydf$steps)
filling <- activitydf %>% group_by(interval) %>% 
      summarise(med = median(steps, na.rm = TRUE))
processdf <- activitydf
processdf[filt, "steps"] <- filling$med[filling$interval == processdf[filt, "interval"]]
```


Histogram of the total number of steps taken each day :
```{r na.totalstep.hist, message = FALSE}
temp <- processdf %>% group_by(date) %>%
      summarise(stepsperday = sum(steps, na.rm = TRUE))
hist(temp$stepsperday,
     xlab = "Steps per day",
     main = "Numbers of days for a certain amount of steps per day")
```


Mean and median of the total number of steps taken per day :
```{r na.mean.median, message = FALSE}
processdf %>% group_by(date) %>%
      summarise(stepsperday = sum(steps, na.rm = TRUE)) %>%
      summarise(mean.stepsperday = mean(stepsperday, na.rm = TRUE),
                median.stepsperday = median(stepsperday, na.rm = TRUE)) %>%
      as.data.frame
```

With this NAs filling strategy :  
* histogram seems identical from the previous  
* the mean value is the same as before (10395 steps)  
* the median value is a little lower (before / after = 9354.23 / 9372.934 steps)  

## Are there differences in activity patterns between weekdays and weekends?

```{r daysweek.plot, fig.height=8, fig.width=10, message=FALSE}
library(lubridate)
library(lattice)
processdf <- processdf %>% mutate(wd = factor(wday(date), labels = c(rep("weekday", 5), rep("weekend", 2))))
temp <- processdf %>% group_by(wd, interval) %>% 
      summarise(avg = mean(steps, na.rm = TRUE))
xyplot(avg ~ interval | wd, data = temp,
       layout = c(1,2), type='l', col = "#53979E", lwd = 2,
       xlab = "5-minutes interval [HoursMinutes]", ylab = "Average number of steps",
       main = "Average number of steps taken of the work days or weekend days by 5-minutes day interval")

```

